---
title: "esm_206_lab05"
author: "Derek Nguyen"
date: "10/29/2019"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

 ### LAb 5 objectives
 
 - learn to parse dates with 'lubridate'
 - get counts of observations with count()
 - uncount() rows
 - one and two-sample t-test
 - create a heatmap with geom_tile()
 
```{r, include = FALSE}
# Read in data and attach packages
library(tidyverse)
library(here)
library(janitor)


lobster_abundance <- read_csv(here::here("data","lobster_abundance.csv"), 
                              na = "-99999") %>% 
  janitor::clean_names()
  
```

Use tidyr::uncount() function to convert our lobster data from frequency format to case format

```{r}
lobster_tidy <- lobster_abundance %>% 
  tidyr::uncount(lobster_count)
```

### Exploratory data visualization

we are only going to consider 'site' as our variable of interest.

```{r}
ggplot(lobster_tidy, aes(x = site, y = size_mm)) + 
  geom_jitter(aes(color = site),
              alpha = 0.5,
              width = 0.2) 

# Histogram

ggplot(lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site),
                 alpha = 0.7) +
  facet_wrap(~site, scales = "free") # scale = free allows for scales to be independent for each site...



ggplot(lobster_tidy, aes(sample = size_mm)) +
  geom_qq() +
  facet_wrap(~site) # facet wrap separates observations by a variable = 'site'

```

#### convert date column to class 'date'

WE'll use the 'lubridate' package to convert to Date format, and then to help us easily parse month & year....

```{r}

lobster_date <- lobster_tidy %>% 
  mutate(
    date_new = lubridate::mdy(date)
  )


```

Now lets parse year and month using lubridate::month() and lubridate::year()

```{r}
lobster_parse_date <- lobster_date %>% 
  mutate(
    obs_month = lubridate::month(date_new, label = TRUE),
    obs_year = lubridate::year(date_new)  #lubridate knows month should be 
  )
```


Now lets find counts of observed lobsters based on different hierarchical groupings:

First, lets count lobsters by year and month:
```{r}
lobster_ym <- lobster_parse_date %>% 
  dplyr::count(obs_year, obs_month)

lobster_y <- lobster_parse_date %>% 
  dplyr::count(obs_year)


lobster_y

lobster_site <- lobster_parse_date %>% 
  dplyr::count(site)

lobster_site
# 
```

If we want to create a summary table that contains stats other than counts by group, it is easier to use group_by() + n().....

```{r}
#only group by site
lobster_summary <- lobster_parse_date %>% 
  group_by(site) %>% 
  summarize(
    mean_size = mean(size_mm, na.rm = TRUE),
    sd_size = sd(size_mm, na.rm = TRUE),
    lobster_number = n() # =multiple summary stats by group, n() is counts within the group
  )

#Tally() = convenient wrapper for summarize that will either call n() or sum(N) 

lobster_summary
```

#### Find Confidence Intervals

Use t-test() function to find confidence intervals for one sample and perform t-tests to compare means of two samples....

```{r}
ivee_lobster <- lobster_tidy %>% 
  filter(site == "IVEE") %>% 
  pull(size_mm) # pull function 

t.test(ivee_lobster)
```

#### Two-sample t-test to compare means

We want to ask: is there a significant difference in lobster lengths at Naples and Mohawk reefs?

We've done our necessary exploratory analyses to determine that a 2-sample t-test for means comparison is appropriate

```{r}

napl_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm)


mohk_sample <- lobster_tidy %>% 
  filter(site =="MOHK") %>% 
  pull(size_mm)


mn_ttest <- t.test(mohk_sample, napl_sample)


mn_ttest
```

There is a significant difference in lobster lengths between Naples and Mohawk Reef. (t(df)) = statistic, p < 0.001, alpha = 0.05).


```{r}
lobster_mn <- lobster_tidy %>% 
  filter(site %in% c("NAPL", "MOHK"))


# ~ means to look in site column, theres two groups, want to do a comparison between two
mn_ttest2 <- t.test(size_mm ~ site, data = lobster_mn)

mn_ttest2

mn_ttest2$p.value 
mn_ttest2$statistic
```

Here is the p-value: `r_mn_ttest$p.value`

Here is the t-statistic: `r mn_ttest2$statistic`


#### Now a HEAT MAP = two discrete values, colors = variable value

```{r}

lobster_ys <- lobster_parse_date %>% 
  count(obs_year, site)


ggplot(data = lobster_ys, aes(x = obs_year, y = site)) +
  geom_tile(aes(fill = n))
```















